\documentclass[letterpaper]{article}
%\documentclass[letterpaper,10pt]{scrartcl}

\usepackage[utf8]{inputenc}
\usepackage[pass,letterpaper]{geometry}

\title{Non-blocking,In-memory Database\thanks{Sponsor: Dr. Dechev}}
\author{Michael McGee, Robert Medina, Neil Moore, Jason Stavrinaky}
\date{9/18/2015}

\pdfinfo{%
  /Title    (Non-blocking, In-memory Database)
  /Author   (Michael McGee, Robert Medina, Neil Moore, Jason Stavrinaky)
  /Creator  (Neil Moore)
  /Producer ()
  /Subject  (Initial Project Description for COP4934)
  /Keywords ()
}

\begin{document}
  \pagenumbering{gobble}
  \maketitle
  \newpage
  \pagenumbering{arabic}

  \section{Introduction}
  This project's goal is to create a database that utilizes advanced multiprocessor techniques and algorithms to fully utilize modern hardware.
  We will be utilizing the Tervel framework that was developed by Dr. Dechev and his lab that implements multiprocessor-friendly memory management
  as well as various lock-free and wait-free data structures. These data structures will permit us to handle many concurrent transactions without 
  waits or locks.
  
  \subsection{Motivations}
  \subsubsection{Michael McGee's Motivation}
  \textless Insert actual text here\textgreater
  \subsubsection{Robert Medina's Motivation}
  \textless Insert actual text here\textgreater
  \subsubsection{Neil Moore's Motivation}
  My motivation for working on this project is to further my knowledge on multiprocessor programming as well as high-performance computing.  
  As hardware vendors put more and more cores into computers, parallelization of algorithms and programs will become much more important than
  single-threaded performance. During my internship, I worked with both GPU compute (OpenCL) and virtualization technologies which exposed me
  to both multiprocessor and system programming. I really enjoyed working on those projects and I think this project will be the next step 
  in my progression as a software engineer and computer scientist.
  \subsubsection{Jason Stavrinaky's Motivation}
  \textless Insert actual text here\textgreater
  
  \section{Requirements}
  Our deliverable requirements are as follows:
  \begin{itemize}
   \item Able to run simple SQL benchmark such as the one bundled into SQLite.
   \item GitHub Page similar to the Tervel framework's.
   \item Source code on the UCF GitHub subdomain.
   \item In-depth documentation on project internals and architecture.
   \item Workshop paper that details new approaches used in database.
  \end{itemize}
  
  \section{Proposed Architecture}
  \textless Insert block diagrams and initial component interaction\textgreater
  
  \section{Financial Concerns}
  Our sponsor is supplying us with a 64-core server that should be sufficient for our development needs. I do not believe any further financial
  support would be required.
  
  \section{Project Milestones (as of 9/??/2015)}
  \textless Insert milestones for project\textgreater

\end{document}
